Red Wine Data Analysis by Raahul Seshadri
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
winedata <- read.csv('wineQualityReds.csv')
```

> **Tip**: Before you create any plots, it is a good idea to provide a short
introduction into the dataset that you are planning to explore. Replace this
quoted text with that general information!

# Univariate Plots Section

We'll go through the summary of the dataset and a few univariate plots to understand the structure of the data.

## Summary of the dataset

```{r echo=FALSE}
summary(winedata)
```

```{r}
nrow(winedata)
ncol(winedata)
```

We observe the following:

1. There are 1599 observations across 13 features.
1. As mentioned on the website, there are no missing values for any of the features.
1. All the fields are numeric in nature.
1. Quality is the only field that's human-defined. Rest of the fields are chemical properties of the wine.

## Distribution of quality

```{r echo=FALSE}
ggplot(aes(x = quality), data = winedata) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:10, limits = c(1, 10))
```

Observations:
1. Quality ranges from 3 to 8.
2. No wine in the dataset has been rated with a quality of 1, 2, 9, and 10.
3. Most of the wine have a quality rating of 5 or 6.
4. The distribution is approximately normal around 5 and 6, although 7 has a greater count than 4.

## Distribution of a few numeric features

The bin-width has to set by looking at the min/max value. I explicitly set the `bins` instead of the `binwidth`, since it is more convenient. Every variable has a different range.

### Fixed acidity

```{r echo=FALSE}
ggplot(aes(x = fixed.acidity), data = winedata) +
  geom_histogram(bins = 30) +
  scale_x_continuous(breaks=)
```

Observations:
1. The data is positively skwewed with mean at 8.32 (as per the statistics we calculated earlier)

### pH

The value of pH ranges from 0 to 14 (15 values). I've chosen 30 as the bin size so that I get a 0.5 resolution on the pH.

```{r echo=FALSE}
ggplot(aes(x = pH), data = winedata) +
  geom_histogram(bins=30) +
  scale_x_continuous(breaks=seq(0, 14, 0.5))
```

Observations:
1. The distribution is again normal centered around 3.31.

### Density

I iterated on the number of bins from 10 to 20 to 40, till I was satisfied with the visual resolution.

```{r echo=FALSE}
ggplot(aes(x = density), data = winedata) +
  geom_histogram(bins=40)
```

Observations:
1. Again, normally distributed around 0.9967

# Univariate Analysis

### What is the structure of your dataset?
There are 1599 observations across 13 features. As mentioned on the website, there are no missing values for any of the features.

All the fields are numeric in nature. Quality is the only field that's human-defined. Rest of the fields are chemical properties of the wine.

All the data examine so far were normally distributed. Between the basic summary of the data and the histograms, I thought boxplots would have added little value.

### What is/are the main feature(s) of interest in your dataset?

The main feature of interest was the "quality", since it was human-determined. There are no categorical variables of interest, so all numeric variables with normal distribution look about the same in a univariate analysis.

To get more information about which numeric features are of importance, we'll need to do a bivariate analysis on the numeric feature and quality. The correlation between numeric features can also be found to detect redundance.

### What other features in the dataset do you think will help support your \
The dataset did mention that the quality is the median of at least 3 evaluations made by wine experts. It'd be interesting to know the difference in the scoring of the same wine between evaluations. Any difference above 1 on either side could be statistically significant and can indicate the presence of bias.

### Did you create any new variables from existing variables in the dataset?
No, I did not.

### Of the features you investigated, were there any unusual distributions? \
I did not transform any of the data. Although the Udacity video mentioned that "normalizing" data is required to apply linear regression, it is incorrect. The data can be in any format, the error has to be normally distributed.

As such, no tidying was done because most of the data were already normally distributed.

# Bivariate Plots Section

## Relationship between various acidity parameters

I see three features on acidity, "fixed.acidity", "volatile.acidity", "citric.acidity". It'd be interesting to know if there are any relationships between them.

```{r echo=FALSE}
ggplot(aes(x = fixed.acidity, y = volatile.acidity), data = winedata) +
  geom_point()
```

```{r echo=FALSE}
ggplot(aes(x = fixed.acidity, y = citric.acid), data = winedata) +
  geom_point()
```

```{r echo=FALSE}
ggplot(aes(x = volatile.acidity, y = citric.acid), data = winedata) +
  geom_point()
```

```{r}
cor.test(winedata$fixed.acidity, winedata$volatile.acidity, method='pearson')
cor.test(winedata$fixed.acidity, winedata$citric.acid, method='pearson')
cor.test(winedata$citric.acid, winedata$volatile.acidity, method='pearson')
```

Observations:
1. Fixed and volatile acidity don't have a strong correlation (pearson coeff of magnitude 0.25)
1. However, fixed and citric show signs of positive correlation (pearson coeff of 0.67)
1. Volatile and citric acid show signs of negative correlation (pearson coeff of -0.55)

## pH and density

```{r echo=FALSE}
ggplot(aes(x = pH, y = density), data = winedata) +
  geom_point()
```

```{r}
cor.test(winedata$pH, winedata$density, method='pearson')
```

Observations:
1. A linear relationship IS NOT observed.

## Relationship to the quality

Let's examine the relationship of a few variables to the quality.

### Fixed acidity vs. quality

```{r echo=FALSE}
ggplot(aes(x = fixed.acidity, y = quality), data = winedata) +
  geom_point()
```

```{r}
cor.test(winedata$fixed.acidity, winedata$quality, method='pearson')
```

Observation:
1. No direct relation between fixed acidity and quality observed (pearson coeff of 0.124)

### Residual sugar vs. quality

```{r echo=FALSE}
ggplot(aes(x = residual.sugar, y = quality), data = winedata) +
  geom_point()
```

```{r}
cor.test(winedata$residual.sugar, winedata$quality, method='pearson')
```

Observation
1. Both quality 3 and 8 wines have lower values of residual sugar. So residual sugar is not alone a strong indicator of quality. (pearson coeff of 0.013)

### Alchohol vs. quality

```{r echo=FALSE}
ggplot(aes(x = alcohol, y = quality), data = winedata) +
  geom_point()
```

```{r}
cor.test(winedata$alcohol, winedata$quality, method='pearson')
```

Observation:
1. We see some trend in the lower bound for alcohol, where lower values can be observed for lower quality of wine.
1. However, the max values for alcohol do not indicate any correlation to the quality.
1. The pearson coefficient is 0.476

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

1. I found no strong correlation between quality and a few features that I examined with no pair exceeding a pearson coefficient of 0.5.
2. Most likely, wine quality would depend on several parameters at once, and can probably be revealed by running linear regression with all the features with quality as the target variable.
3. PCA can be used to find feature importances.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

1. There was a strong positive correlation between fixed.acidity and citric.acid.
2. There was a strong negative correlation between citric.acid and volatile.acidity.

### What was the strongest relationship you found?

The relationship between fixed.acidity and citric.acid was the strongest, with a pearson coefficient of 0.67.

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!